---
title: "Intro to Strings"
author: "Mike J"
date: "4/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Things you can and can't do with strings in R

```{r nostrings }
e <- 'Mixed quotes" 
print(e)

f <- 'Single quote ' inside single quote'
print(f)

g <- "Double quotes " inside double quotes"
print(g)

```

## Including Plots

You can also embed plots, for example:

```{r cando, echo=FALSE}
string1 <- "This is a string"
string1


string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
string2
```
```{r bad}
ex3 <- "This "is" totally unacceptable"

ex4 <- 'This 'is' absolutely wrong'
```
```{r example}
example[1] <- "first"
example

length(example)
```

#The basic syntax for format function is −

format(x, digits, nsmall, scientific, width, justify = c("left", "right", "centre", "none"))
#Following is the description of the parameters used −

x is the vector input.

digits is the total number of digits displayed.

nsmall is the minimum number of digits to the right of the decimal point.

scientific is set to TRUE to display scientific notation.

width indicates the minimum width to be displayed by padding blanks in the beginning.

justify is the display of the string to left, right or center.


```{r formats}
# Total number of digits displayed. Last digit rounded off.
result <- format(23.123456789, digits = 9)
print(result)

# Display numbers in scientific notation.
result <- format(c(6, 13.14521), scientific = TRUE)
print(result)

# The minimum number of digits to the right of the decimal point.
result <- format(23.47, nsmall = 5)
print(result)

# Format treats everything as a string.
result <- format(6)
print(result)

# Numbers are padded with blank in the beginning for width.
result <- format(13.7, width = 6)
print(result)
class(result)

# Left justify strings.
result <- format("Hello", width = 8, justify = "l")
print(result)

# Justfy string with center.
result <- format("Hello", width = 8, justify = "c")
print(result)
```

To change strings to all upper case or all lower case

The basic syntax for toupper() & tolower() function is −

```{r cases}
x<-"heLLo"

toupper(x)
tolower(x)
nchar(x)


```


substring(x,first,last)
Following is the description of the parameters used −

x is the character vector input.

first is the position of the first character to be extracted.

last is the position of the last character to be extracted.

```{r substring}
# Extract characters from 5th to 7th position.
result <- substring("Extract", 5, 7)
print(result)

x <- c("d", "a", "c", "abba")  

#where is the character located in the vector
grep("a", x)

base <- "Wow that was cool."
sub("cool", "AWESOMAZING!", base)
```

paste: Concatenate Strings

```{r paste}
a <- "Hello"
b <- 'How'
c <- "are you? "

print(paste(a,b,c))


print(paste(a,b,c, sep = "-"))

print(paste(a,b,c, sep = "", collapse = ""))
```


Count the character within a string  nchar()

```{r ct}

#With the help of this function, we can count the characters. 
str <- "Big Data at DataFlair"
 nchar(str)
 
```
sprintf()
This function makes of the formatting commands that are styled after C.

```{r spr}
 
sprintf("%s scored %.2f percent", "Matthew", 72.3)

sprintf("%2$f scored %1$s percent", "Matthew", 72.3)
```

```{r subst}
x1 <- "Get excited" 
substr(x1, start = 5, stop = 13)


```

```{r dataflair}
#Author DataFlair
num <- "12345678"
substr(num, 4, 5)
substr(num, 5, 7)

x <- c("d", "a", "c", "abba")      # Create example character vector
t=grep("a", x)
t

# Creating a sequence from 1 to 9
x <- 1:9

# Calling cat() function
cat(x, sep =" + ")
cat("\n")
cat(x, sep =" / ")


```


```{r replacev}
#a input vector 
df<-"R is a collaborative project with many contributors"

#replaces the string 
sub('R','R language',df)



```

```{r replace}
#creating a data frame
df<-data.frame(Column_1 = c('Florida','Germany','Georgia','Geniva','Istanbul','NewZealand','Australia'), Column_2=c(1,2,3,4,5,6,7))

#data frame
df

#replacing the G character with A
sub('G','A',df)
```


Stringr

```{r stringr }
#install.packages("stringr")
library(stringr)

#count length
str_length(c("a", "R for data science", NA))

#concatenate
str_c("x", "y")

#concatenate
str_c("x", "y", "z")

#concatenate with a separator
str_c("x", "y", sep = ", ")

#concatenate with a prefix or suffix
str_c("prefix-", c("a", "b", "c"), "-suffix")

#separate out string values
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)

# negative numbers count backwards from end
str_sub(x, -3, -1)

```
```{r strup }
str_to_upper(c("i", "ı"))

#change to different language Turish
str_to_upper(c("i", "ı"), locale = "tr")

```

Split strings

```{r splits}
fruits <- c(
  "apples and oranges and pears and bananas",
  "pineapples and mangos and guavas"
)
#remove the string and
str_split(fruits, " and ")

```

```{r}
#Author DataFlair

#split into pieces

str = "Splitting sentence into words"
strsplit(str, " ")

#replace string with new text

x <- c("why", "video", "cross", "extra", "deal", "authority")
str_replace(x, "[aeiou]", "?")

```

